---
title: "Introduction to Content Validity Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Content Validity Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(contentvalidity)
```

## Overview

The `contentvalidity` package provides comprehensive tools for calculating content validity metrics used in scale development and validation. This vignette introduces the main concepts and demonstrates how to use the package.

## What is Content Validity?

Content validity refers to the extent to which a measurement instrument adequately covers the construct it is designed to measure. When developing a new questionnaire, survey, or assessment tool, researchers typically ask a panel of subject matter experts to evaluate whether each item is relevant to the construct being measured.

## Key Metrics

### Content Validity Index (CVI)

The **Content Validity Index (CVI)** is the most widely used metric for quantifying content validity. It can be calculated at two levels:

- **I-CVI (Item-level CVI)**: The proportion of experts who rate an item as relevant
- **S-CVI (Scale-level CVI)**: Aggregated measure for the entire instrument
  - **S-CVI/Ave**: Mean of all I-CVI values
  - **S-CVI/UA**: Proportion of items with Universal Agreement

### Universal Agreement (UA)

**Universal Agreement** indicates whether ALL experts rated an item as relevant. Items with UA = 1 have perfect agreement among experts.

### Content Validity Ratio (CVR)

**Lawshe's CVR** provides a statistical basis for content validity:

$$CVR = \frac{n_e - N/2}{N/2}$$

Where $n_e$ is the number of experts rating the item as essential and $N$ is the total number of experts.

## Basic Usage

### Preparing Your Data

Your data should be organized with:
- **Rows** = Experts (raters)
- **Columns** = Items being evaluated
- **Values** = Numeric ratings

```{r example-data}
# Example: 5 experts rating 4 items on a 4-point scale
ratings <- data.frame(
  item1 = c(4, 4, 3, 4, 3),  # Highly relevant
  item2 = c(4, 4, 4, 4, 4),  # Perfect agreement
  item3 = c(3, 2, 2, 3, 2),  # Mixed ratings
  item4 = c(4, 3, 4, 4, 3)   # Good agreement
)
ratings
```

### Running the Analysis

Use the `cvi()` function for a comprehensive analysis:

```{r run-analysis}
result <- cvi(ratings)
print(result)
```

### Detailed Summary

For more detailed output with interpretation:

```{r summary}
summary(result)
```

### Visualizing Results

Plot I-CVI values with threshold lines:

```{r plot, fig.width=7, fig.height=5}
plot(result)
```

## Individual Metrics

You can calculate individual metrics directly:

```{r individual-metrics}
# Item-level CVI
icvi(ratings)

# Scale-level CVI (Average method)
scvi_ave(ratings)

# Scale-level CVI (Universal Agreement)
scvi_ua(ratings)

# Universal Agreement
ua(ratings)

# Content Validity Ratio
cvr(ratings)
```

## Flexible Rating Scales

The package supports multiple rating scale formats:

### 4-Point Scale (Default)

```{r scale-4point}
# 1 = not relevant, 2 = somewhat, 3 = quite, 4 = highly relevant
# Values 3-4 are coded as "relevant"
result_4pt <- cvi(ratings, scale = "4point")
```

### 3-Point Scale

```{r scale-3point}
ratings_3pt <- data.frame(
  item1 = c(3, 3, 2, 3),
  item2 = c(2, 2, 2, 1)
)
# Values 2-3 are coded as "relevant"
result_3pt <- cvi(ratings_3pt, scale = "3point")
```

### Binary Scale

```{r scale-binary}
ratings_binary <- data.frame(
  item1 = c(1, 1, 1, 0),
  item2 = c(1, 1, 1, 1)
)
result_binary <- cvi(ratings_binary, scale = "binary")
```

### Custom Scale

```{r scale-custom}
# 5-point scale where 4-5 are "relevant"
ratings_5pt <- data.frame(
  item1 = c(5, 5, 4, 5),
  item2 = c(4, 4, 4, 5),
  item3 = c(3, 2, 4, 3)
)
result_custom <- cvi(ratings_5pt, scale = "custom", relevant_values = c(4, 5))
```

## Interpretation Guidelines

Based on established literature, use these thresholds:

| Metric | Threshold | Interpretation |
|--------|-----------|----------------|
| I-CVI | >= 0.78 | Acceptable (Lynn, 1986) |
| S-CVI/Ave | >= 0.90 | Excellent (Polit & Beck, 2006) |
| S-CVI/UA | >= 0.80 | Acceptable |

### CVR Critical Values

For CVR, significance depends on panel size. Use `cvr_critical()` to get the threshold:

```{r cvr-critical}
# Get critical value for 10 experts
cvr_critical(10)

# Check if items meet threshold
n_experts <- 5
critical_val <- cvr_critical(n_experts)
item_cvr <- cvr(ratings)
item_cvr >= critical_val
```

## Working with Expert Statistics

The analysis also provides per-expert statistics:

```{r expert-stats}
result <- cvi(ratings)
result$expert_stats
```

View in summary:

```{r expert-summary}
summary(result, show_experts = TRUE)
```

## Accessing Raw Binary Data

If you need the transformed binary ratings:

```{r binary-data}
result$binary_data
```

Or transform directly:

```{r transform-direct}
transform_ratings(ratings, scale = "4point")
```

## References

- Lynn, M. R. (1986). Determination and quantification of content validity. *Nursing Research*, 35(6), 382-385.

- Polit, D. F., & Beck, C. T. (2006). The content validity index: Are you sure you know what's being reported? Critique and recommendations. *Research in Nursing & Health*, 29(5), 489-497.

- Lawshe, C. H. (1975). A quantitative approach to content validity. *Personnel Psychology*, 28(4), 563-575.
